\documentclass{acmart}

\usepackage[utf8]{inputenc}
\usepackage{url}

\parindent 0cm

\begin{document}

\onecolumn

\pagestyle{empty}
\begin{center}
{\bfseries \Huge Chairs' Welcome}
\end{center}

\bigskip

%\large

It is our great pleasure to welcome you to FARM 2017, the 5th ACM
SIGPLAN International Workshop on Functional Art, Music, Modelling and
Design!

\smallskip

The ACM SIGPLAN International Workshop on Functional Art, Music,
Modelling and Design (FARM) gathers together people who are harnessing
functional techniques in the pursuit of creativity and expression.

\smallskip

Functional Programming has emerged as a mainstream software
development paradigm, and its artistic and creative use is booming. A
growing number of software toolkits, frameworks and environments for
art, music and design now employ functional programming languages and
techniques. FARM is a forum for exploration and critical evaluation of
these developments, for example to consider potential benefits of
greater consistency, tersity, and closer mapping to a problem domain.

\smallskip

By gathering researchers, practitioners, artists, designers and anyone
interested, FARM aims to develop this interdisciplinary field. Work in
this challenging area has already led to new language designs,
abstraction techniques, and execution models, as well as the use of
old ideas in new, novel, and surprising ways.

\smallskip

After its first edition held in Boston in 2013, its second edition
held in Gothenburg in 2014, the third in Vancouver 2015, and the
fourth in Nara in 2016, this year's FARM will take place in Oxford,
UK on September 9.

\smallskip

As has become the norm, FARM 2017 is dominated by contributions on
computer music: Halley Young presents a categorical grammar for
music. Edward Lilley also presents a new representation for music
notation.  Chris Ford will talk about representing African polyphony
and polyrhythm.  Yann Orlarey, Stéphane Letz, Dominique Fober, and
Romain Michon present the FAUST DSL for real-time signal processing
and synthesis.  Thomas Murphy will present a sound-synthesis system in
Haskell based on SuperCollider.

\smallskip

However, not all is music at the 2017 FARM: A representation for
mathematics, presented by Joseph Corneli, Raymond Puzio, Ursula Martin, Dave Murray-Rust, Alison Pease, and Gabriela Rino Nesin breaks new
ground for FARM.  Simon Archipoff and David Janin present their
approach for unified media programming, accompanied by a demo of a 3D
animation language.  Ivan Perez presents a graphic adventure library
and engine.  Finally, Stian Veum Møllersen describes Ait, a
concatenative language for creative programming.

\smallskip

In 2014, 2015, and 2016 FARM also organized an evening of performances on
the last day of ICFP.  This year continues what is becoming a
tradition, with a line-up of exciting computer-music performances
assembled by Alex McLean, headlined by Alexandra C\'ardenas.

\smallskip

Further information about FARM is available on our website,
\url{functional-art.org}.

\smallskip

We would like to thank ACM and the organizers of ICFP 2017 for their
support, especially general chair Jeremy Gibbons
as well as the workshop chairs, David Christiansen and Andres Löh. Thanks also
to the members of the FARM Program Committee, and to everyone who
submitted papers, demos, and performance proposals. And, of course, we
are grateful to all of the participants. We hope that you will enjoy
the workshop as much as we know we will.

\medskip

Welcome to the FARM!

\bigskip
%\bigskip




\hskip -.25cm 
\begin{tabular}[t]{l}
{\bf Michael Sperber}\\
Workshop Chair\\
Active Group GmbH
\end{tabular}
\hfill
\begin{tabular}[t]{l}
 {\bf Jean Bresson}\\
Program Committee Chair\\
IRCAM
\end{tabular}
\hfill
\begin{tabular}[t]{l}
{\bf Alex McLean}\\
  Performance Chair\\
  Deutsches Museum, Munich
\end{tabular}
\hfill
\begin{tabular}[t]{l}
  {\bf Mark Santolucito}\\
  Publicity Chair\\
  Yale University
\end{tabular}

\vspace{2mm}

\newpage
\begin{center}
{\bfseries \Huge FARM 2017 Workshop Organization}
\end{center}

\bigskip

\large
\begin{center}
\begin{tabular}{rl}

{\bf Workshop Chair:} & Michael Sperber (\emph{Active Group GmbH})\\
\\

{\bf Program Chair:} & Jean Bresson (\emph{UMR STMS IRCAM/CNRS/UPMC Sorbonne Universités})\\
  \\

  {\bf Performance Chair:} & Alex McLean (\emph{Deutsches Museum, Munich})\\
  \\

{\bf Publicity Chair:} & Mark Santolucito (\emph{Yale University})\\

\\
  
  {\bf Program Committee:} &
                             Renick Bell (\emph{Tama Art University, Tokyo})\\
  &Beatrice Berard, LIP6 (\emph{Universite Pierre et Marie Curie})\\
  &  Jean Bresson (\emph{UMR STMS IRCAM/CNRS/UPMC Sorbonne Universités})\\
  &  Richard Gabriel (\emph{IBM Research})\\
   & Jean-Louis Giavitto (\emph{UMR STMS IRCAM/CNRS/UPMC Sorbonne Universités})\\
   & Stéphane Letz (\emph{Grame})\\
   & José Pedro Magalhães (\emph{Standard Chartered Bank})\\
   & Ivan Perez Dominguez (\emph{University of Nottingham})\\
   & Donya Quick (\emph{Southern Methodist University})\\
    &Christophe Rhodes (\emph{Goldsmiths, University of London})\\
   & Andrew Sorensen (\emph{Australian National University})\\
   & Michael Sperber (\emph{Active Group GmbH})\\
   & Charlotte Truchet (\emph{LINA, UMR 6241, Université de Nantes})
\end{tabular}
\end{center}

\vspace{2mm}

\end{document}
